---
title: "Assignment 05"
author: "SABIHA MHATARNAIK"
date: "4/25/2021"
output: html_document
---
```{r}
library(dplyr)
library(cluster)
library(fpc)
```

```{r}
setwd("~/Desktop/Machine Learning/Assignment_05")
```
#Importing dataset in R
```{r}
cereals.df<-read.csv("Cereals.csv")
```
#Eliminating unwanted variables from the dataset
```{r}
cereals.df <-select(cereals.df,-c('mfr','type'))
```
#Giving the row number as name and eliminating the "name" from the dataset to avoid data duplication
```{r}
rownames(cereals.df)<-cereals.df$name
cereals.df$name=NULL
```
#Data Normalization
```{r}
cereals.df<-as.data.frame(scale(cereals.df))
```
#Data Preprocessing
```{r}
sum(is.na(cereals.df))
cereals.df<-na.omit(cereals.df)
```
#Dissimilarity matrix
```{r}
d <-dist(cereals.df,method="euclidean")
```
#Q1Hierarchical clustering and Plotting the dendrogram

#Apply hierarchical clustering to the data using Euclidean distance to the normalized measurements. Use Agnes to compare the clustering from single linkage, complete linkage, average linkage, and Ward. Choose the best method.
```{r}
hc_single <- hclust(d,method="single")#Single linkage: computes the minimum distance between clusters before merging them
plot(hc_single,cex=0.6,hang=-1)

hc_complete <- hclust(d,method="complete")#Complete linkage: computes the maximum distance between clusters before merging them.
plot(hc_complete,cex=0.6,hang=-1)

hc_average <- hclust(d,method="average")#Average linkage: computes the average distance between clusters before merging them.
plot(hc_average,cex=0.6,hang=-1)

hc_ward <- hclust(d,method="ward.D")# Ward’s (minimum variance) criterion: minimizes the total within-cluster variance and find the pair of clusters that leads to minimum increase in total within-cluster variance after merging.
plot(hc_ward,cex=0.6,hang=-1)
```


#Compute with Agnes and different linkage methods 

#Compare Agglomerative coefficients
```{r}
hc_Single <- agnes(d,method="single")
print(hc_Single$ac)

hc_Complete <- agnes(d,method="complete")
print(hc_Complete$ac)

hc_Average <- agnes(d,method="average")
print(hc_Average$ac)

hc_Ward <- agnes(d,method="ward")
print(hc_Ward$ac)
```

#The Best Linkage method is Ward with Agglomerative coefficient of 0.9049881.

#Plot Dendrogram
```{r}
pltree(hc_Ward,cex=0.6,hang=-1,main="Dendrogram of Ward")
```

#2Q Choosing the number of clusters

```{r}
d <-dist(cereals.df,method="euclidean")
hc_ward <- hclust(d,method="ward.D")
plot(hc_ward,cex=0.6)
rect.hclust(hc_ward,k=3,border=1:3)
clust<-cutree(hc_ward,k=3)
```

#The optimal number of clusters is 3 by the longest distance and the partition.

#3Q Stability of clusters
#Comment on the structure of the clusters and on their stability. Hint: To check stability,partition the data and see how well clusters formed based on one part apply to the other part. To do this:

● Cluster partition A

● Use the cluster centroids from A to assign each record in partition B (each record is assigned to the cluster with the closest centroid).

● Assess how consistent the cluster assignments are compared to the
assignments based on all the data.
```{r}
set.seed(123)
kmeans_clust <- kmeans(cereals.df,3,nstart=10)
km_data <- kmeans_clust$cluster
cereal_cluster <- as.data.frame(cbind(kmeans_clust$cluster,cereals.df))
colnames(cereal_cluster)[1]<-"cluster"
print_clusters <-function(labels,k)
{
  for(i in 1:3)
  {
    print(paste("cluster", i))
    print(cereal_cluster[labels==i,c("cluster","calories" ,"protein" , "fat" ,     "sodium"  , "fiber"  , "carbo"  ,  "sugars"  , "potass"  , "vitamins" ,"shelf" ,   "weight"  , "cups"  ,   "rating"  )])
    
  }
}
groups <- cutree(hc_ward, k=3)
print_clusters(groups,3)
kbest.p<- 3       
cboot.hclust <- clusterboot(cereals.df,clustermethod=hclustCBI,method="ward.D", k=kbest.p)
groups<-cboot.hclust$result$partition 
cboot.hclust$bootmean 
which.max(cboot.hclust$bootmean)
```
#Cluster 3 has highest stability(cluster stability) amongst all.

#4Q Cluster with healthy Cereal and High nutrition value
#The elementary public schools would like to choose a set of cereals to include in their daily cafeterias. Every day a different cereal is offered, but all cereals should support a healthy diet. For this goal, you are requested to find a cluster of “healthy cereals.”Should the data be normalized? If not, how should they be used in the cluster analysis?
```{r}
mydata <- na.omit(cereals.df)
rating <- cbind(mydata,clust)
rating[rating$clust==1,]
rating[rating$clust==2,]
rating[rating$clust==3,]
```
#Cluster 1 has high protein and fibre content compared to others .Hence,it can be said that cluster 1 possess high nutritional value and has healthy cereals .

#Also cluster 1 consists of fruits ,nuts,wheat which can be considered as a healthy cereal.

#Yes,the data should be normalized as we are uing distance as a measure of performance.