---
output:
  word_document: default
  html_document: default
---
#Analysis of dataset in R
##By Sabiha Mhatranaik

#Load Dataset packages
```{r}
library(readxl)
```
#import the excel file into the global environment
```{r}
Dataset_Media_ <- read_excel("Machine Learning/Dataset_Media -.xls")
```
#head function gives the first 6 rows
```{r}
head(Dataset_Media_)
```

#tail function gives the last 6 rows 
```{r}
tail(Dataset_Media_)
```
#str function will structure the data 
```{r}
str(Dataset_Media_)
```
#summary function will give the summary of the data 
```{r}
summary(Dataset_Media_)
```
```{r}
plot(Dataset_Media_)
```
#X-Y plot for two quantitative variables(Scatterplot)
```{r}
plot(Dataset_Media_$Audience_Size,Dataset_Media_$Preference)
```
#Add some Options
```{r}
plot(Dataset_Media_$Audience_Size,Dataset_Media_$Preference,pch=20,col="#cc0000",main="Example of Media",xlab="Size of the Audience",ylab="Preference")
```
#Plotting a quantitative variable Audience_Size
```{r}
plot(Dataset_Media_$Audience_Size)
```
#Lineplot,Histogram,Boxplot
```{r}
plot(Dataset_Media_$Audience_Size,type="l",main="Example of Media",xlab="Size of the Audience",ylab="Preference")
hist(Dataset_Media_$Audience_Size)
boxplot(Dataset_Media_$Audience_Size)
```
#Working with functions
```{r}
min(Dataset_Media_$Audience_Size)
max(Dataset_Media_$Audience_Size)
range(Dataset_Media_$Audience_Size)
mean(Dataset_Media_$Audience_Size)
median(Dataset_Media_$Audience_Size)
```
#The End


